CALVIS: chest, waist and pelvis circumference from 3D human body meshes as ground truth for deep learning

### 2020-03-04 PaperRMFI2019/Response_to_Reviewers.tex

\documentclass{article}
\usepackage{a4wide}
\title{Response to the Reviewer's Comments}
\author{Yansel Gonzalez Tejeda \and	Helmut Mayer}
\begin{document}

\maketitle

We thank the reviewers for identifying the weaknesses in our paper and 
providing us the opportunity to strengthen our research prior to publication.

The comments of the reviewers have been very useful, for they
made us aware that some major points we were trying to convey
remained obscure even to well-informed readers.

Following up
on the reviewers' suggestions, we have restructured the paper
by changing the order of some sections, by putting more emphasis
on motivation, and by adding clarifications at crucial spots.

In addition, we have also corrected all factual errors and
minor infelicities of presentation that the reviewers brought
to our attention. Below we indicate for each individual comment
how we have dealt with it.

\section*{Reviewer 1}
\begin{verbatim}
> 1)English and equations in the current paper are terrible. The authors need
> to improve the spelling, structure and presentation of the paper.
\end{verbatim}

\begin{verbatim}
> 2) The authors used many abbreviation words without providing any explanation.
\end{verbatim}

For all abbreviation we add the corresponding meaning. For example, Skinned 
Multi-Person Linear (SMPL),

\begin{verbatim}
> 3) There are several typo errors.
> For example:
> pg 3 :
> The SCAPE model[3] opened wide possibilities in the field. ??? which field ?
\end{verbatim}

We added the clarification: "in the field of human shape estimation".

\begin{verbatim}
> has been strongly influence --> influenced
> no human body dimensions are computed of estimated.--> from the estimated
> model
> do not require that level of detail. --> such level of detail
\end{verbatim}

 We have corrected these errors.

\begin{verbatim}
> 4) The authors used the SMPL generative human body model, developed in 2015
> to synthesize 3D human meshes.
\end{verbatim}

Yes, we do. Here we are not able to grasp the intention of the 
reviewer.

\begin{verbatim}
> 5) pg 13, it is not accurate to say "The automatic annotation process took
> approximately 1 h 47 min in an enhanced modern personal computer." What are
> the characteristics of the computer ?
\end{verbatim}

\begin{verbatim}
> 6) The description of the bloc diagram of the CNN Calvis-Net depicted in fig
> 6, page 12, is missed; The proposed CNN architecture is not convincing. No
> detail has been provided about the relation and the functioning of the
> weight, the convolutional and the FC layers.
\end{verbatim}

\begin{verbatim}
> 7) The results are far from being convincible. Experimental results are with
> limited 3D articulated meshes. It would be better if some quantified
> indicators are used (Accuracy assessment).
\end{verbatim}

We take the point, and we think our new version is much clearer.

\section*{Reviewer 2}
\begin{verbatim}
> 1). It is not clear how 8 images can be used to train CNN with ground trut
> of 3 measures.
\end{verbatim}
\begin{verbatim}
> 2). It is not clear how novel the proposed measurement calculation is.
\end{verbatim}
\begin{verbatim}
> 3). there is no comparison with existing methods
\end{verbatim}

We read the objection of the reviewer as a complaint about the
story line. Our restructuring should have remedied this.

We now emphasize the real reason why we want the reduction axioms:
as a tool for {\em compositional analysis} of the epistemic effects of
many different types of information-carrying event. It is actually
surprising that this is possible. We see this compositional analysis
as the heart of dynamic epistemic logic, not just a negotiable
'luxury' for some cases. As a side-effect, such axioms
'reduce' a complete dynamic epistemic language to its static
epistemic base -- but this is a consequence of the analysis,
not the main motivation. Of course, as we now also stress, this
reduction does have another interesting interpretation by itself:
the static language must be rich enough to 'pre-encode' the
effects of the relevenat class of informational events. This
is a requirement on optimal language design which is also known
from other areas of logic, such as conditional logic vis-a-vis
belief revision. Finally, and we may have given the wrong
impression that *this* was our main point: the reduction
analysis tends to make for simpler completeness proofs,
and some direct borrowing of known properties of the
static base into the full dynamic system.

What we also find interesting is how this methodology puts us on
the track of new epistemic operators that have not been isolated
as important before, even though they were there just below the
surface. In particular, relativized common knowledge seems a
new notion of independent interest.

\section*{Reviewer 3}
\begin{verbatim}
> Overall, the paper is very well written. However, the main contributions are
> not clear.
\end{verbatim}

\begin{verbatim}
> The segmentation method is also too constrained as it requires the
> shapes to be a pre-defined position.
\end{verbatim}

\begin{verbatim}
> The paper should also try and compare with other existing segmentation
> algorithms. This is a very well studied topic, see for example:
> https://scholar.google.com.au/scholar?
> q=3d+mesh+segmentation&hl=en&as_sdt=0&as_vis=1&oi=scholart
\end{verbatim}

While we appreciate the reviewer recommendation, it has to be said that we do 
not focus on a comparative analysis of segmentation algorithms. Nevertheless, 
we do think that we could perform such as analysis on future works.

\section*{Reviewer 4}
\begin{verbatim}
> Thank you for sharing the generated database.
\end{verbatim}

We are glad to contribute and thank the reviewer for this encouraging, 
motivational and 
constructive comment.

\section*{Reviewer 5}
\begin{verbatim}
> The main contribution of this paper is to computing the human body dimensions
> based on the geometric methods and meanwhile to testing the CNN-based
> regression method. However, I have the following concerns:
> 1. The experimental details related to the CNN-based method are missing. The
> detailed hyper-parameters used for CNN training and testing are missing.
\end{verbatim}

The reviewer is right. We add the experimental details. See also address to 
reviewer 5

\begin{verbatim}
> 2. There is no comparisons with other related methods. The error of 12mm
> achieved in this paper has no reference score.
\end{verbatim}

\begin{verbatim}
> 3. Overall, I think the problem of 3D human shape measurement is a important
> task, and the geometric method used in this paper has a certain contribution
> to this task.
\end{verbatim}

We are glad to contribute and thank the reviewer for this encouraging, 
motivational and 
constructive comment.

\begin{verbatim}
> 4. The presentation of this paper is not smooth.
\end{verbatim}

The new presentation makes our intentions clearer, we think.
The completeness proof of the logic of epistemic action models as
provided by Baltag, Moss and Solecki is complicated. Our method of
reduction axioms allows us to `lift' the elegant PDL completeness
proof to the logic of epistemic action models. And again: the
reduction axioms that we find are interesting in their own right
as we are after compositional analysis of epistemic effects of events
observed in groups of agents, and earlier dynamic epistemic logic were
simply unable to provide that methodology for common knowledge.

The `point' of LCC is to show that `really' epistemic PDL, contrary to
what the designs of intricate logics of communication and knowledge
suggest, is all that is needed for a rich logic of communication and
change. So if anything, this is a new vindication of PDL, rather than
a criticism of it.

It has to be said, though, that we do not provide a knock-down argument
why this full system is {\em needed} for our style of analysis. We have
not been able to find anything weaker that works, but our list of
open questions at the end does ask if other weaker 'solutions' exist.

\section*{Reviewer 7}
\begin{verbatim} 
> Very interesting work with a recent bibliographic study, nevertheless the
> experimentation part lacks comparative study with the existing one and
> especially in term of algorithmic complexity
\end{verbatim}

We hope that our new version is much more focussed.

The author is right on one further point. We have chosen to add
events that change the world to the standard DEL-framework, since
we feel that this increases the scope with little effort, while
still operating with the same methodology. But it is clear that
this is orthogonal to our main points. We have considered taking
it out (but that seemed a pity), or to put it into a separate
section: but that would lead to too much duplication. Our solution
now is to leave the story combined, while pointing out to the reader
who so prefers that a 'change-free' reading of our main notions and
results can be obtained by merely omitting all our technical talk
about substitutions.

\begin{verbatim}
> As far as style is concerned, the work "perspicuous" appears three
> times on the first page; once in the abstract, and twice in the
> second paragraph. While perspicuous is a delicious word, it is rare
> enough that seeing it three times makes one frown.
\end{verbatim}

We have rephrased things using a wider repertoire of adjectives.

\begin{verbatim}
> Also, typo on p.3,
> first paragraph of 2.1: "PLaza" instead of "Plaza".
\end{verbatim}

Corrected.

\begin{verbatim} 
>>REFEREE REPORT 2<<
> List of typos, corrections and comments:
>
> 1. One of the key definitions ...
\end{verbatim}

The referee is correct that this definition was wrong.

The following change solves the problem. We now take the transitive
closure rather than the *reflexive* transitive closure in the
definition. We have adapted the rest of the text including the
axioms accordingly, and now also the remark on page 5 is correct.

\end{document}

### 2020-03-04 PaperRMFI2019/Paper_ID_25_CameraReady.tex
update

\begin{document}
%
\title{CALVIS: chest, waist and pelvis circumference from 3D human body meshes	
as ground truth for deep learning}
%
\titlerunning{Chest, waist and pelvis circumference from 3D human body meshes}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Author1\orcidID{0000-0000-0000-0000} \and
Author2\orcidID{0000-0000-0000-0000}}
%
\authorrunning{Author1 and Author2}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Institute1}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In this paper we present CALVIS, a method to calculate \textbf{C}hest, 
w\textbf{A}ist and pe\textbf{LVIS} circumference from 3D human body meshes. Our 
motivation is to use 
this data as ground truth for training convolutional neural networks 
(CNN). Previous work had used the large scale CAESAR dataset or determined 
these anthropometrical measurements $\textit{manually}$ from a person 
or human 3D body meshes. 
Unfortunately, acquiring these data is a cost and 
time consuming endeavor. In contrast, our method can be used on 
3D meshes automatically. We synthesize eight human body meshes and 
apply 
CALVIS to calculate chest, waist and pelvis circumference. We evaluate the 
results qualitatively and observe that the measurements can indeed be used 
to estimate the shape of a person. We then asses the 
plausibility of our approach by generating
ground truth with CALVIS to train a small CNN. After
having trained the network with 
our data, we achieve competitive validation error.
Furthermore, we make the implementation of CALVIS publicly available to 
advance the field.

@@

In the past decade a noticeably amount of researchers have employed this 
dataset to investigate human shape estimation. Because the measurement 
acquiring process is resource intensive and requires large amount of human and 
material resources, this type of studies are rare and the data they offer is 
expensive. Therefore, it is important to explore alternative methods where 
human body measurements derived from real data can be obtained for 
investigation.

3D human body generative models offers such an alternative. In the line of previous work, we start by synthesizing 3D human body meshes using the SMPL \cite{Loper.2015} generative 
model. Unlike other works, once we have 
the 3D meshes we compute chest, waist and pelvis circumference. The next 
step after obtaining the 
measurements is to use a camera model to render
images. Finally, in possession of this ground truth we can 
input these images to the learning algorithm and train with the human body dimensions as supervision signal.

@@

%-------------------------------------------------------------------------
\subsection{Human Dimensions Estimation from Images.}
Using metrology techniques \cite{ams.2008.BenAbdelkaderY08} and mixture of experts model \cite{Sigal.2008} human body dimensions like height, arm span and even weight had been estimated from images. Chest and waist size had been considered as well \cite{Guan.2013}. These investigations used human dimensions for validation (ground truth) that were directly recovered from individuals. In contrary, we calculate them  
consistently from 3D meshes.
Another significant research direction studies shape as (non necessarily meaningful) parameters of some model. For example \cite{Bogo:ECCV:2016} claim describing the first method 
to 
estimate pose and shape from a single image, while \cite{DBLP:Lassner2017} estimates shape from 91 keypoints using decision 
forests. Shape is understood as the 
parameters of the SMPL model but not human dimensions.

@@

\subsection{Human Body Data Synthesis.}
The SCAPE model\cite{Anguelov.2005} opened wide possibilities in the field. It 
provided the scientific community with a method to build realistic 3D human 
body meshes of different shapes in different poses. In order to synthesize a 
human body mesh a template body must be deformed. The model exhibits, however, an important limitation, namely the last 
step on the pipeline is an expensive optimization problem. Other researchers had used variation of the SCAPE model (S-SCAPE \cite{Pishchulin.2017}) to synthesize human bodies but focused on people detection.

After some attempts on improving the human models quality, for example, to make 
more 
easily capturing the correlation between body shape and pose 
\cite{HaslerSSRS09}, or to better understand body dimensions (SPRING model 
\cite{Yang.2014}), Loper et al., 2015 \cite{Loper.2015} developed the SMPL generative 
human body model. In this work we synthesize 3D human meshes using this 
model. We will briefly describe the model in subsection \ref{subsec:smpl_model}.

Our approach to synthesize images from 3D meshes has been strongly influence by the recent publication of the large scale dataset SURREAL \cite{varol17_surreal}. This work uses the SMPL model 
to generate images 
from humans with random backgrounds. However, no human body dimensions are computed of estimated.

More recently, a human model with added hands and face had been presented \cite{Joo.2018}. We do not use this more complex model 
because the human dimensions we estimate do not require that level 
of detail.

@@

\subsection{SMPL Generative Model of the Human Body}\label{subsec:smpl_model}
In this work we synthesize 3D human meshes using the SMPL
model \cite{Loper.2015}. The generative nature of our approach establishes this 
model as starting point (and not the 3D mesh). Nevertheless, our method is 
flexible enough to begin the pipeline with a 3D mesh. In that case, an SMPL 
model can be fitted to the volumetric data, using the method described by 
Varol et al., 2018 \cite{varol18_bodynet} for example.

@@

Once we have a segmented mesh, we search the dimensions within 
these regions. The body dimensions we want to calculate are 
geodesic on the mesh. Next subsection discuss how.

@@

Figure \ref{fig:qualitative_eval} shows the calculated dimensions on eight 
subjects, male subjects no. 1-4 and female subjects no. 5-8. The black paths 
represent chest, waist and pelvis circumference (from 
top to bottom) as defined by equations \ref{eq:cc}-\ref{eq:pc}, respectively. 
We slightly rotate the meshes and add transparency for visualization purposes. 
Note that our method is able to estimate automatically the human body 
dimensions for all 
subjects. The dimensions are indeed estimated at locations where domain experts regularly identify them.

@@

\subsubsection{Experimental setup.} First, we assemble our Calvis-CMU Dataset. We synthesize 3400 grayscale images (1700 of each gender) of human body meshes generated with parameters from the CMU dataset using Blender. To increase the realism of these images we 
adopt an indoor lighting model with one light source and white background. This makes sense because it resembles the way 
how a person is standing in a room when she attends to the tailor for her 
body dimensions to be taken.
Subsequently, we annotate the CMU dataset with CALVIS. The 
automatic annotation process took approximately $1\,h\, 47\, min$ in an 
enhanced modern personal computer.

@@

Figure \ref{fig:calvis_net} shows our prototypical CNN Calvis-Net where the 
input is a synthetic image from a 3D human mesh and the output are three body 
dimensions. Implemented with pytorch \cite{paszke2017automatic}. We train the network for 20 epochs and perform 3-fold cross-validation  with skorch. We achieve mean absolute validation error of 12 mm.

>>>>

\begin{document}
%
\title{CALVIS: chest, waist and pelvis circumference from 3D human body meshes	
as ground truth for deep learning}
%
\titlerunning{Chest, waist and pelvis circumference from 3D human body meshes}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Yansel Gonzalez-Tejeda\orcidID{0000-0003-1002-3815} \and
	Helmut Mayer\orcidID{0000-0002-2428-0962}}
%
\authorrunning{Gonzalez-Tejeda and Mayer}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Paris Lodron University of 
	Salzburg, Kapitelgasse 4-6, 5020 Salzburg, Austria
	\url{https://www.uni-salzburg.at}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In this paper we present CALVIS, a method to calculate \textbf{C}hest, 
w\textbf{A}ist and pe\textbf{LVIS} circumference from 3D human body meshes. Our 
motivation is to use 
this data as ground truth for training convolutional neural networks 
(CNN). Previous work had used the large scale CAESAR dataset or determined 
these anthropometrical measurements $\textit{manually}$ from a person 
or human 3D body meshes. 
Unfortunately, acquiring these data is a cost and 
time consuming endeavor. In contrast, our method can be used on 
3D meshes automatically. We synthesize eight human body meshes and 
apply 
CALVIS to calculate chest, waist and pelvis circumference. We evaluate the 
results qualitatively and observe that the measurements can indeed be used 
to estimate the shape of a person. We then asses the 
plausibility of our approach by generating
ground truth with CALVIS to train a small CNN. After
having trained the network with 
our data, we achieve competitive validation error.
Furthermore, we make the implementation of CALVIS publicly available to 
advance the field.

@@

In the past decade a noticeable number of researchers have employed this 
dataset to investigate human shape estimation. Because the measurement 
acquiring process is resource intensive and requires large amount of human and 
material resources, this type of studies are rare and the data they offer is 
expensive. Therefore, it is important to explore alternative methods where 
human body measurements derived from real data can be obtained for 
investigation.

3D human body generative models offers such an alternative. In the line of 
previous work, we start by synthesizing 3D human body meshes using the Skinned 
Multi-Person Linear (SMPL) \cite{Loper.2015} generative 
model. Unlike other works, once we have 
the 3D meshes we compute chest, waist and pelvis circumference. The next 
step after obtaining the 
measurements is to use a camera model to render
images. Finally, in possession of this ground truth we can 
input these images to the learning algorithm and train with the human body dimensions as supervision signal.

@@

%-------------------------------------------------------------------------
\subsection{Human Dimensions Estimation from Images.}
Using metrology techniques \cite{ams.2008.BenAbdelkaderY08} and mixture of 
experts model \cite{Sigal.2008} human body dimensions like height, arm span and 
even weight had been estimated from images. Chest and waist size had been 
considered as well \cite{Guan.2013}. These investigations used human dimensions 
for validation (ground truth) that were directly recovered from individuals. In 
contrast, we calculate them  
consistently from 3D meshes.
Another significant research direction studies shape as (non necessarily meaningful) parameters of some model. For example \cite{Bogo:ECCV:2016} claim describing the first method 
to 
estimate pose and shape from a single image, while \cite{DBLP:Lassner2017} estimates shape from 91 keypoints using decision 
forests. Shape is understood as the 
parameters of the SMPL model but not human dimensions.

@@

\subsection{Human Body Data Synthesis.}
The Shape Completion and Animation of People (SCAPE) model\cite{Anguelov.2005} 
opened wide possibilities in the field of 
human shape estimation. It 
provided the scientific community with a method to build realistic 3D human 
body meshes of different shapes in different poses. In order to synthesize a 
human body mesh a template body must be deformed. The model exhibits, however, an important limitation, namely the last 
step on the pipeline is an expensive optimization problem. Other researchers had used variation of the SCAPE model (S-SCAPE \cite{Pishchulin.2017}) to synthesize human bodies but focused on people detection.

After some attempts on improving the human models quality, for example, to make 
more 
easily capturing the correlation between body shape and pose 
\cite{HaslerSSRS09}, or to better understand body dimensions (Semantic 
Parametric Reshaping of Human Body Models - SPRING model 
\cite{Yang.2014}), Loper et al., 2015 \cite{Loper.2015} developed the SMPL generative 
human body model. In this work we synthesize 3D human meshes using this 
model. We will briefly describe the model in subsection \ref{subsec:smpl_model}.

Our approach to synthesize images from 3D meshes has been influenced 
by the recent publication of the large scale dataset SURREAL 
\cite{varol17_surreal}. This work uses the SMPL model 
to generate images 
from humans with random backgrounds. However, no human body dimensions are 
computed or estimated.

More recently, a human model with added hands and face had been presented \cite{Joo.2018}. We do not use this more complex model 
because the human dimensions we estimate do not require such level 
of detail.

@@

\subsection{SMPL Generative Model of the Human Body}\label{subsec:smpl_model}
In this work we synthesize 3D human meshes using the Skinned Multi-Person 
Linear (SMPL) model \cite{Loper.2015}. The generative nature of our approach 
establishes this 
model as starting point (and not the 3D mesh). Nevertheless, our method is 
flexible enough to begin the pipeline with a 3D mesh. In that case, an SMPL 
model can be fitted to the volumetric data, using the method described by 
Varol et al., 2018 \cite{varol18_bodynet} for example.

@@

Once we have a segmented mesh, we search the dimensions within 
these regions. Next subsection discuss how to calculate geodesic on the mesh to 
establish a search strategy.

@@

Figure \ref{fig:qualitative_eval} shows the calculated dimensions on eight 
subjects, male subjects no. 1-4 and female subjects no. 5-8. The black paths 
represent chest, waist and pelvis circumference (from 
top to bottom) as defined by equations \ref{eq:cc}-\ref{eq:pc}, respectively. 
We slightly rotate the meshes and add transparency for visualization purposes. 
Note that our method is able to calculate automatically the human body 
dimensions for all 
subjects. The dimensions are indeed calculated at locations where domain 
experts regularly identify them.

@@

\subsubsection{Experimental setup.} First, we assemble our Calvis-CMU Dataset. 
Using Blender we render 3400 grayscale images (1700 of each gender) of 
human body meshes generated with parameters from the CMU dataset. To increase 
the realism of these images we 
adopt an indoor lighting model with one light source and white background. This makes sense because it resembles the way 
how a person is standing in a room when she attends to the tailor for her 
body dimensions to be taken.
Subsequently, we annotate the CMU dataset with CALVIS. The 
automatic annotation process took approximately $1\,h\, 47\, min$ in an 
enhanced modern personal computer.

@@

Figure \ref{fig:calvis_net} shows our prototypical CNN Calvis-Net implemented 
with pytorch \cite{paszke2017automatic}. The 
input is a synthetic grayscale image from a 3D human mesh of size $\mathtt{200 
x 200 x 
1}$. The input layer is followed by a convolutional layer $\mathtt{CL}_1$ with 
$\mathtt{3}$ 2D-kernels of 
size $\mathtt{5 x 5}$ and max pooling $\mathtt{MP}$ over a $\mathtt{2 x 2}$ 
volume. The network continues then with the rectified linear activation function
(ReLU). The last layer (fully connected $\mathtt{FC}$) regresses directly three 
body 
dimensions: chest, waist and pelvis circumference. We train the network for 20 
epochs and perform 3-fold 
cross-validation, achieving mean absolute validation error of 12 
mm.
